---
title: "Logistic Regression - Framingham Heart Study"
author: "Casandra Philipson"
date: "February 26, 2016"
output: html_document
---

##Overview: The Framingham Heart Study  
_Evaluating risk factors to save lives_  
  
Original project was inspired by FDR's rising blood pressure during presidency. Framingham, MA was chosen as a location with a stable popilation to be used as patients during the study. William Kannell and Roy Dawber led the Framingham Heart Study wherein 5,209 patients aged 30-59 enrolled in 1948 and were given an exam every 2 years.  
  
The __goal__ of this project is to: _predict 10-year risk for Coronary Heart Disease_.  
  
###Risk Factors: Analytics to Prevent Heart Disease  
Identifying appropriate _risk factors_ is the key to succecssful prediction of coronary heart disease (CHD).  
  
```{r}
framingham <- read.csv("framingham.csv")
str(framingham)
```

__Demographic Risk Factors__  
  
* sex
* age
* education
  

__Behavioral Risk Factors__  
  
* currentSmoker  
* cigsPerDay  
  

__Medical History Risk Factors__  
  
* BPmeds  
* prevalentStroke  
* prevalentHyp  
* diabetes  
  
__Examination Risk Factors__  
  
* totChol  
* sysBP  
* diaBP  
* BMI  
* heartRate  
* glucose  
  

###Logistic Regression Model
__1__. Randomly split patients into training and testing sets.  
```{r}
library(caTools)

# Randomly split the data into training and testing sets
set.seed(1000)
split <- sample.split(framingham$TenYearCHD, SplitRatio = 0.65)

# Split up the data using subset
train <- subset(framingham, split==TRUE)
test <- subset(framingham, split==FALSE)

```

__2__. Use logistic regression on training set to predict whether or not a patient experienced CHD within 10 years of the first examination.  

```{r}
# Logistic Regression Model
framinghamLog <- glm(TenYearCHD ~ ., data = train, family = binomial)
summary(framinghamLog)
```
Evaluating all variables as part of the model allows us to identify the following risk factors as significant or important:  
_male, age, cigsPerDay, prevalentStroke, prevalentHyp, totChol, sysBP, glucose_  
  
  
__3__. Evaluate the predictive power on the test set.
```{r}
# Predictions on the test set
predictTest <- predict(framinghamLog, type = "response", newdata = test)

# Confusion matrix with threshold of 0.5
table(test$TenYearCHD, predictTest > 0.5)
```

The accuracy of our model [Overall accuracy = (TN + TP) / N] is:
```{r}
(1069+11)/(1069+6+187+11)
```

This is barely better than the baseline accuracy:
```{r}
(1069+6)/(1069+6+187+11) 
```

__4__. Evaluate Model Strength on Test Set
```{r}
suppressMessages(library(ROCR))
ROCRpred <- prediction(predictTest, test$TenYearCHD)
as.numeric(performance(ROCRpred, "auc")@y.values)
```

###Conclusions
  
* Model rarely predicts 10-year CHD risk above 50% (accuracy very near baseline of always predicting no CHD)  
* Model _can_ differentiate low-risk from high-risk patients (AUC = 0.74)  
* Some significant variables suggest interventions:  
    + Smoking, Cholesterol, Systolic BP, Glucose  